Supplementary Materials for Manuscript of MADA Project

### 2023-05-05 products/manuscript/Manuscript.qmd
Final Project Update

Women aged 40-74 years were reported having mammogram as 84.5%, 85.3%, 87.8%, and 87.5% during years' 2000, 2005, 2010, and 2015 respectively. Hispanic women were found having lower mammogram screening rates (80.0%) compared to non-Hispanic women (87.5%). All bivariate and multivariate relationships between mammogram status and hypothesized predictors (year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status) were found to be significantly associated with mammogram screening. There was a higher log odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold. Alcohol and smoke users were found having low log odds of having mammograms compared to never alcohol and smoke users.

@@

The univariate relationship between the mammogram screening and each hypothesized predictors was assessed using logistic regression. Multiple logistic regression was also performed for determining the multivariate effect on mammogram screening. Model selection process was performed and the final model model was selected based on ROC AUC (Area under the Receiver Operating Characteristic curve) value. The model evaluation was performed using 5-fold cross-validation process repeated for 5 times. All analysis were performed using R (version 4.2.1) and cross-validation was performed using the tidymodels package. The code and data for reproducing results are available on [Github](https://github.com/shiwanisapkota/shiwanisapkota-MADA-project).

@@

We aimed to explore the bivariate and multivariate relationship between mammogram screening and other hypothesized predictors i.e., year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status. Logistic regression was used for examining such relationships. In bivariate relationship (Supplementary Tables 3-12), we found statistically significant relationship between mammogram screening and hispanic status of women and found the log odds of having mammogram screening for Hispanic women as -0.56 compared to non-Hispanic women. The log odds of having mammogram was found as 0.25 in year 2015 compared to that of 2000. We also found that the log odds of having mammogram increased with age, and education. We found the log odds of having mammogram screening for women with income below the poverty threshold as -0.76 compared to women with income at or above the poverty threshold. The log odds of having mammogram for married was 0.67 compared to never married women. Women having the usual medical place had the log odds of having mammogram of 1.47 compared to those having no usual medical place. Similarly, Table 2 shows the multivariate relationship between mammogram screening and all other predictors of interests. All predictors of interests were found to be significant predictors of mammogram.

Table 3 shows the model performance comparison which compared the models' performance including various numbers of predictors. The best model performance was found for the full model including all predictors of interests (AIC: 16981 and RMSE: 0.32). Similarly, based on model performance comparison, second best model was identified as the model including demographic predictors (age, hispanic status, income, education, marital status, and region of residence) that had AIC of 17764.1 and RMSE of 0.33.

@@

# Loading the `model_performance_comparison.rds` table which was created during the model fitting analysis process
result_table3 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
kable(result_table3 %>% select(Name, AIC, BIC, RMSE),  align = "c", format.args = list(digits = 3)) %>%
  row_spec(row = 0, color = "white") %>% kable_styling(full_width = T, bootstrap_options = "striped", font_size = 10)
```

@@

In this study, we examined the mammography screening status of women aged 40 to 74 years and their relationship with the hypothesized predictors: year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status. During years 2000, 2005, 2010, and 2015, 84.5%, 85.3%, 87.8%, and 87.5% of women aged 40-74 years reported having mammogram. Though there was not much increment in mammography rate from 2010 to 2015 but there was some mammography rate increment compared to 2000 and 2005. Such increment in mammography rate could be due to people's increased awareness and education, improved access to healthcare, advanced in technology, and changing mammography recommendation guidelines. Our findings support previous research as Hispanic women having lower mammogram screening rates (80.0%) compared to non-Hispanic women (87.5%) (@selvin_breast_2003-1). All bivariate and multivariate relationships between mammogram status and predictors were found to be significant hence, all hypothesized predictors were found to have strong association with mammogram screening. As suggested in previous studies (@katz_breast_2000), we found higher log odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold. Alcohol and smoke users were found having low log odds of having mammograms compared to never alcohol and smoke users. Previous studies have also suggested that people engaging in high risk behaviors are less likely to comply with cancer screening guidelines (@holm_health_1999). This finding could be of great importance as health practitioners could identify such individuals and recommend appropriate cancer screening services. In full model, we found education and age as the predictors having highest significant association with mammogram screening. In that case, intervention programs should emphasize on educating women and make uneducated women aware of such cancer screening services.

@@

This study confirms that there exists some association between mammogram screening and socio-economic status of Hispanic/non-Hispanic women during different years as all the hypothesized predictors were found to be significantly associated with mammogram status. We found age, education, income, and smoking/alcohol behavior as the great predictors for having mammogram screening. Further research needs to be done to dive more into these hypothesized predictors to see their absolute and relative effects on mammogram screening. Furthermore, given the lower overall breast cancer screening rates among Hispanic women, increased education and awareness should be among top priorities for developing appropriate population-based interventions for increasing the use of mammography screening.

>>>>

Women aged 40-74 years were reported having mammogram as 84.5%, 85.3%, 87.8%, and 87.5% during years' 2000, 2005, 2010, and 2015 respectively. Hispanic women were found having lower mammogram screening rates (80.0%) compared to non-Hispanic women (87.5%). All bivariate and multivariate relationships between mammogram status and hypothesized predictors (year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status) were found to be significantly associated with mammogram screening. There was a higher odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold. More educated women had higher odds of having mammogram compared to low or even non-educated women. Current smokers were found having lower odds of having mammograms compared to never smokers.

@@

The univariate relationship between the mammogram screening and each hypothesized predictors was assessed using logistic regression. Multiple logistic regression was also performed for determining the multivariate effect on mammogram screening. Model selection process was performed and the final model was selected based on ROC AUC (Area under the Receiver Operating Characteristic curve) value. The model evaluation was performed using 5-fold cross-validation process repeated for 5 times. All analysis were performed using R (version 4.2.1) and cross-validation was performed using the tidymodels package. The code and data for reproducing results are available on [Github](https://github.com/shiwanisapkota/shiwanisapkota-MADA-project).

@@

We aimed to explore the bivariate and multivariate relationship between mammogram screening and other hypothesized predictors i.e., year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status. Logistic regression was used for examining such relationships. In bivariate relationship (Supplementary Tables 3-12), we found statistically significant relationship between mammogram screening and hispanic status of women and found the odds of having mammogram screening for Hispanic women as 0.57 compared to non-Hispanic women. The odds of having mammogram was found as 1.28 in year 2015 compared to that of 2000. We also found that the odds of having mammogram increased with age, and education. We found the odds of having mammogram screening for women with income below the poverty threshold as 0.47 compared to women with income at or above the poverty threshold. The odds of having mammogram for married women was 1.96 compared to never married women. Women having the usual medical place had the odds of having mammogram as 4.34 compared to those having no usual medical place. Similarly, Table 2 shows the multivariate relationship between mammogram screening and all other predictors of interests. All predictors of interests were found to be significant predictors of mammogram.

Table 3 shows the model performance comparison which compared the models' performance including various numbers of predictors. The best model performance was found for the full model including all predictors of interests as it had the lowest AIC value of 16981 and RMSE value of 0.32. Similarly, based on model performance comparison, second best model was identified as the model including demographic predictors (age, hispanic status, income, education, marital status, and region of residence) that had AIC of 17764.1 and RMSE of 0.33.

@@

# Loading the `model_performance_comparison.rds` table which was created during the model fitting analysis process
result_table3 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
kable(result_table3 %>% select(Name, AIC, BIC, RMSE),  align = "c", format.args = list(digits = 3)) %>% 
  row_spec(row = 0, color = "white") %>% kable_styling(full_width = T, bootstrap_options = "striped", font_size = 10)
```
*Note: The given model name includes mammogram status as a response variable and the following variables as predictors as given in brackets: log_fit_year (year), log_fit_hispanicstatus (hispanic status of women), log_fit_age (age), log_fit_income (income), log_fit_education (education), log_fit_maritalstatus (marital status), log_fit_regionresidence (region of residence), log_fit_insurancestatus (insurance status), log_fit_healthstatus (health status), log_fit_usualmedicalcarestatus (usual place for medical care), log_fit_smokingstatus (smoking status), log_fit_alcoholstatus (alcohol status), log_fit_diabetesstatus (diabetes status), log_fit_fullmodel (all predictors of interests), log_fit_year_age_hispanic (year, age, and hispanic status), log_fit_demographic (age, hispanic status, income, education, marital status, and region of residence), log_fit_medical (insurance status, health status, usual place for medical care, and diabetes status), and log_fit_smoking_alcohol (smoking status and alcohol status).

@@

In this study, we examined the mammography screening status of women aged 40 to 74 years and their relationship with the hypothesized predictors: year, hispanic status, age, income, education, marital status, region of residence, insurance status, health status, usual medical care status, smoking status, alcohol status, and diabetes status. During years 2000, 2005, 2010, and 2015, 84.5%, 85.3%, 87.8%, and 87.5% of women aged 40-74 years reported having mammogram. Though there was not much increment in mammography rate from 2010 to 2015 but there was some mammography rate increment compared to 2000 and 2005. Such increment in mammography rate could be due to people's increased awareness and education, improved access to healthcare, advanced in technology, and changing mammography recommendation guidelines. Our findings support previous research as Hispanic women having lower mammogram screening rates (80.0%) compared to non-Hispanic women (87.5%) (@selvin_breast_2003-1). All bivariate and multivariate relationships between mammogram status and predictors were found to be significant hence, all hypothesized predictors were found to have strong association with mammogram screening. As suggested in previous studies (@katz_breast_2000), we found higher odds of having mammogram among women with income at or above the poverty threshold as opposed to those below the poverty threshold. We also found older aged and highly educated women to have higher odds of having mammogram screening compared to young and low educated women. Current smokers were found having lower odds of having mammograms compared to people who never smoked. Previous studies have also suggested that people engaging in high risk behaviors are less likely to comply with cancer screening guidelines (@holm_health_1999). This finding could be of great importance as health practitioners could identify such individuals and recommend appropriate cancer screening services. In full model, we found education and age as the predictors having highest significant association with mammogram screening. In that case, intervention programs should emphasize on educating women and make uneducated women aware of such cancer screening services.

@@

This study confirms that there exists some association between mammogram screening and socio-economic status of Hispanic/non-Hispanic women during different years as all the hypothesized predictors were found to be significantly associated with mammogram status. We found age, education, income, and smoking behavior as the great predictors for having mammogram screening. Further research needs to be done to dive more into these hypothesized predictors to see their absolute and relative effects on mammogram screening. Furthermore, given the lower overall breast cancer screening rates among Hispanic women, increased education and awareness should be among top priorities for developing appropriate population-based interventions for increasing the use of mammography screening.

### 2023-05-05 products/supplement/Supplementary_Material.qmd
Final Project Update

---
title: "Supplement to Manuscript for MADA Project"
subtitle: "Supplement Materials"
author: Shiwani Sapkota
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(knitr)
library(ipumsr)
library(readr)
library(skimr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for chi-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
```

@@

# Model Evaluation

# Supplementary Table 20
```{r, message = FALSE}
#| echo: FALSE
 # Loading coefficients table using mammogram status as the outcome and other variables of interests as the predictors using model evaluation
result_table20 <- readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds"))
result_table20
```

# Supplementary Table 21
```{r, message = FALSE}
#| echo: FALSE
 # Loading coefficients table using mammogram status as the outcome and hispanic status as the main predictor using model evaluation
result_table21 <- readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds"))
result_table21
```

# Supplementary Table 22

```{r, message = FALSE}
#| echo: FALSE
 # Loading coefficients and penalty data during lasso model process
result_table22 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds"))
result_table22
```

>>>>

---
title: "Supplementary Materials for Manuscript of MADA Project"
subtitle: "Supplement Materials"
author: Shiwani Sapkota
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(knitr)
library(ipumsr)
library(readr)
library(skimr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for chi-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
library(here)
```

@@
